---
title: "Consumer WTP for Carbon Offsets"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(data.table)
library(dplyr)
library(stringr)
library(gtools)
library(gmnl)
library(mlogit)
library(ordPens)
library(nnet)
library(colorBlindness)
library(openxlsx)
library(ggplot2)
rm(list = ls())
load("/home/yufenlai/Documents/others/carbon_footprint/output/ce_est.RData")
dt_demo <- read.xlsx("/home/yufenlai/Documents/others/carbon_footprint/tables/tbl_demo.xlsx")

### define a summary function for nnet
summary_multinom <- 
    function(model) {
        s <- summary(model)
        coef_mat <- t(s$coefficients)
        se_mat <- t(s$standard.errors)
        rownames(coef_mat) <- gsub("X", "", rownames(coef_mat))
        rownames(se_mat) <- gsub("X", "", rownames(se_mat))
        
        tbl <-
            lapply(
                1:ncol(coef_mat),
                function(j) {
                    coef <- coef_mat[, j]
                    se <- se_mat[, j]
                    z <- coef / se
                    pval <- 2 * (1 - pnorm(abs(coef / se)))
                    tbl <-
                        data.frame(
                            coef = coef,
                            se = se,
                            z = z,
                            p_val = pval
                        )
                    return(tbl)
                }
            )
        
        names(tbl) <- paste0("Class_", 2:(length(tbl) + 1))
        return(tbl)
    }
```

# Summary Statistics for respondents
```{r summary_stats, echo=FALSE}
dt_demo %>%
    mutate(
        Percentage = round(Percentage, 2)
    )
```

# Opt-out Rank by Information Treatment
```{r choice_prob, echo=FALSE}
tbl_opt_out_rank <-
    data %>%
    mutate(
        opt_out_rank = case_when(
            alt %in% c(5, 6) ~ 1,
            alt %in% c(1, 3) ~ 2,
            alt %in% c(2, 4) ~ 3
        )
    ) %>%
    filter(
        y == 1
    ) %>%
    group_by(
        framing_effect,
        scenario
    ) %>%
    summarise(
        opt_out_rank = mean(opt_out_rank)
    )

# plot opt_out_rank for each scenario by framing_effect
tbl_opt_out_rank %>%
    ggplot(
        aes(
            x = as.factor(scenario),
            y = opt_out_rank,
            fill = as.factor(framing_effect)
        )
    ) +
    geom_bar(
        stat = "identity",
        position = "dodge"
    ) +
    labs(
        x = "Choice Scenario",
        y = "Choice Rank of Opt-Out",
        fill = "Framing Type"
    ) +
    theme_classic() +
    # colorblind friendly colors
    scale_fill_manual(values = Blue2Gray8Steps) +
    # minimum y = 1.9, maximum y = 2.1
    coord_cartesian(ylim = c(2.2, 2.5))
```

The above table presents the rank of opt-out for each chocie scenario by information treatment.
Lower the rank, higher the preference for opt-out.
The summary shows that Met Office has a poitive effect on the preference for opt-out, and equivalently, the endorsement has a negative effect on the preference for carbon offset.

# Summary of PCA
## for Q9
Factor loadings for Q9:
```{r pca_q9, echo=FALSE}
pc_fit_q9_star$pca
```
Importance of components:
```{r pca_q9_importance, echo=FALSE}
summary(pc_fit_q9_star$pca)
```

## for Q10
Factor loadings for Q10:
```{r pca_q10, echo=FALSE}
pc_fit_q10_star$pca
```
Importance of components:
```{r pca_q10_importance, echo=FALSE}
summary(pc_fit_q10_star$pca)
```

## for Q12
Factor loadings for Q12:
```{r pca_q12, echo=FALSE}
pc_fit_q12_star$pca
```
Importance of components:
```{r pca_q12_importance, echo=FALSE}
summary(pc_fit_q12_star$pca)
```

# Basic Logit Model
## Basic Logit Model Coefficients
```{r basic_logit, echo=FALSE}
options(width = 300)
summary(logit_bsc)
```

## Basic Logit Model Willingness to Pay
```{r basic_logit_wtp, echo=FALSE}
logit_bsc$coefficients["price"] <- logit_bsc$coefficients["price"] * -1
wtp.gmnl(logit_bsc, wrt = "price")
```

# Mixed Logit Model
## Mixed Logit Model Coefficients
```{r mixed_logit, echo=FALSE}
summary(mlogit_bsc)
```

## Mixed Logit Model Willingness to Pay
```{r mixed_logit_wtp, echo=FALSE}
mlogit_bsc$coefficients["price"] <- mlogit_bsc$coefficients["price"] * -1
wtp.gmnl(mlogit_bsc, wrt = "price")
```

# mixed logit + co2 consumption
## mixed logit + co2 consumption Coefficients
```{r mixed_logit_co2, echo=FALSE}
summary(mlogit_ctrl_e)
```

Note that the coefficient for "I.co2_value" measures the utility for opt-out conditional on co2 consumption.
This coefficient is positive means lower baseline utility for carbon offsetting.

## mixed logit + co2 consumption Willingness to Pay
```{r mixed_logit_co2_wtp, echo=FALSE}
mlogit_ctrl_e$coefficients["price"] <- mlogit_ctrl_e$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_e, wrt = "price")
```

# mixed logit + framing effect
## mixed logit + framing effect Coefficients
```{r mixed_logit_framing, echo=FALSE}
summary(mlogit_ctrl_f)
```

## mixed logit + framing effect Willingness to Pay
```{r mixed_logit_framing_wtp, echo=FALSE}
mlogit_ctrl_f$coefficients["price"] <- mlogit_ctrl_f$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_f, wrt = "price")
```

# mixed logit + co2 consumption + framing effect
## mixed logit + co2 consumption + framing effect Coefficients
```{r mixed_logit_co2_f, echo=FALSE}
summary(mlogit_ctrl_ef)
```

The effect of co2 consumption is statistically significant and negative.
However, the effect of framing effect is largely not statistically significant,
and Met Office has a adverse effect on the preference for carbon offsetting and only shift the overall preference, not the preference for carbon offsetting features.

## mixed logit + co2 consumption + framing effect Willingness to Pay
```{r mixed_logit_co2_f_wtp, echo=FALSE}
mlogit_ctrl_ef$coefficients["price"] <- mlogit_ctrl_ef$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_ef, wrt = "price")
```

# mixed logit + 1 PCA for Q9 and Q10 + reduced demographic controls
Here we included the first component for Q9 and Q10.
Q9 was about respondents' attitude towards carbon offsetting,
and Q10 was about respondents' attitude towards climate change.

## mixed logit + 1 PCA for Q9 and Q10 + reduced demographic controls Coefficients
```{r mixed_logit_pca_q9_q10, echo=FALSE}
summary(mlogit_ctrl_p1_910_rd)
```

## mixed logit + 1 PCA for Q9 and Q10 + reduced demographic controls Willingness to Pay
```{r mixed_logit_pca_wtp_q9_q10, echo=FALSE}
mlogit_ctrl_p1_910_rd$coefficients["price"] <- mlogit_ctrl_p1_910_rd$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_910_rd, wrt = "price")
```

# mixed logit + 1 PCA for Q9 and Q10 + full demographic controls
## mixed logit + 1 PCA for Q9 and Q10 + full demographic controls Coefficients
```{r mixed_logit_pca_full_q9_q10, echo=FALSE}
summary(mlogit_ctrl_p1_910_d)
```

## mixed logit + 1 PCA for Q9 and Q10 + full demographic controls Willingness to Pay
```{r mixed_logit_pca_full_wtp_q9_q10, echo=FALSE}
mlogit_ctrl_p1_910_d$coefficients["price"] <- mlogit_ctrl_p1_910_d$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_910_d, wrt = "price")
```

# mixed logit + 1 PCA for Q9 and Q12 + reduced demographic controls
## mixed logit + 1 PCA for Q9 and Q12 + reduced demographic controls coefficients
```{r mixed_logit_pca_q9_q12, echo=FALSE}
summary(mlogit_ctrl_p1_912_rd)
```

## mixed logit + 1 PCA for Q9 and Q12 + reduced demographic controls Willingness to Pay
```{r mixed_logit_pca_wtp_q9_q12, echo=FALSE}
mlogit_ctrl_p1_912_rd$coefficients["price"] <- mlogit_ctrl_p1_912_rd$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_912_rd, wrt = "price")
```

# mixed logit + 1 PCA for Q9 and Q12 + full demographic controls
## mixed logit + 1 PCA for Q9 and Q12 + full demographic controls coefficients
```{r mixed_logit_pca_full_q9_q12, echo=FALSE}
summary(mlogit_ctrl_p1_912_d)
```

## mixed logit + 1 PCA for Q9 and Q12 + full demographic controls Willingness to Pay
```{r mixed_logit_pca_full_wtp_q9_q12, echo=FALSE}
mlogit_ctrl_p1_912_d$coefficients["price"] <- mlogit_ctrl_p1_912_d$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_912_d, wrt = "price")
```

# mixed logit + 1 PCA for Q12 only + reduced demographic controls
## mixed logit + 1 PCA for Q12 only + reduced demographic controls coefficients
```{r mixed_logit_pca_q12, echo=FALSE}
summary(mlogit_ctrl_p1_12_rd)
```

## mixed logit + 1 PCA for Q12 only + reduced demographic controls Willingness to Pay
```{r mixed_logit_pca_wtp_q12, echo=FALSE}
mlogit_ctrl_p1_12_rd$coefficients["price"] <- mlogit_ctrl_p1_12_rd$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_12_rd, wrt = "price")
```

## mixed logit + 1 PCA for Q12 only + full demographic controls Willingness to Pay
```{r mixed_logit_pca_full_wtp_q12, echo=FALSE}
mlogit_ctrl_p1_12_d$coefficients["price"] <- mlogit_ctrl_p1_12_d$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_p1_12_d, wrt = "price")
```

# mixed logit + co2 consumption + framing effect + 2 PCA for Q9 and Q10
We included the first two coponents for Q9 and Q10.
Q9 was about respondents' attitude towards carbon offsetting,
and Q10 was about respondents' attitude towards climate change.

## mixed logit + co2 consumption + framing effect + PCA Coefficients
```{r mixed_logit_co2_f_pca_q9_q10, echo=FALSE}
summary(mlogit_ctrl_efp2_910)
```

## mixed logit + co2 consumption + framing effect + PCA Willingness to Pay
```{r mixed_logit_co2_f_pca_wtp_q9_q10, echo=FALSE}
mlogit_ctrl_efp2_910$coefficients["price"] <- mlogit_ctrl_efp2_910$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_efp2_910, wrt = "price")
```

# mixed logit + co2 consumption + framing effect + 2 PCA for Q9 and Q12
## mixed logit + co2 consumption + framing effect + PCA Coefficients
```{r mixed_logit_co2_f_pca_q9_q12, echo=FALSE}
summary(mlogit_ctrl_efp2_912)
```

## mixed logit + co2 consumption + framing effect + PCA Willingness to Pay
```{r mixed_logit_co2_f_pca_wtp_q9_q12, echo=FALSE}
mlogit_ctrl_efp2_912$coefficients["price"] <- mlogit_ctrl_efp2_912$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_efp2_912, wrt = "price")
```

# mixed logit + co2 consumption + framing effect + 2 PCA for Q12 only
## mixed logit + co2 consumption + framing effect + PCA Coefficients
```{r mixed_logit_co2_f_pca_q12, echo=FALSE}
summary(mlogit_ctrl_efp2_12)
```

## mixed logit + co2 consumption + framing effect + PCA Willingness to Pay
```{r mixed_logit_co2_f_pca_wtp_q12, echo=FALSE}
mlogit_ctrl_efp2_12$coefficients["price"] <- mlogit_ctrl_efp2_12$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_efp2_12, wrt = "price")
```

# mixed logit + co2 consumption + framing effect + 1 PCA for Q12 only
## mixed logit + co2 consumption + framing effect + PCA Coefficients
```{r mixed_logit_co2_f_pca_q12_1, echo=FALSE}
summary(mlogit_ctrl_efp1_12)
```

## mixed logit + co2 consumption + framing effect + PCA Willingness to Pay
```{r mixed_logit_co2_f_pca_wtp_q12_1, echo=FALSE}
mlogit_ctrl_efp1_12$coefficients["price"] <- mlogit_ctrl_efp1_12$coefficients["price"] * -1
wtp <- wtp.gmnl(mlogit_ctrl_efp1_12, wrt = "price")
wtp
```

### WTP Plot, Main Effects and Interavtive Effects with Co2 Consumption and Framing Effect
```{r wtp_plot, echo=FALSE}
# plot WTP for framing effect and co2 consumption
wtp_tbl <- wtp[, c(1, 2)] %>%
    as.data.frame() %>%
    mutate(
        lb = Estimate - 1.96 * `Std. Error`,
        ub = Estimate + 1.96 * `Std. Error`
    )
wtp_tbl$var <- rownames(wtp_tbl)
wtp_tbl <- 
    wtp_tbl %>%
    filter(
        grepl("co2_value|framing_effect", var) &
            !grepl("I.", var)
    ) %>%
    mutate(
        var = 
        str_replace_all(
            var,
            "framing_effect|project_|_value",
            ""
        )
    ) %>%
    mutate(
        var = str_replace_all(var, "_", " ")
    ) %>%
    mutate(
        var = str_replace_all(var, "\\.", " * ")
    ) %>%
    mutate(
        var = str_replace_all(var, "co2", "CO2")
    )

wtp_tbl$var <- factor(wtp_tbl$var, levels = wtp_tbl$var)

# main effects
wtp %>%
    as.data.frame() %>%
    mutate(
        lb = Estimate - 1.96 * `Std. Error`,
        ub = Estimate + 1.96 * `Std. Error`
    ) %>%
    mutate(
        var = rownames(wtp)
    ) %>%
    filter(
        var %in% c(
            "location_EU",
            "location_UK",
            "certificate_NGO",
            "certificate_UK",
            "project_renewable",
            "project_landfill",
            "project_manure"
        )
    ) %>%
    mutate(
        var = str_replace_all(var, "_", " ")
    ) %>%
    mutate(
        var = factor(var, levels = unique(var))
    ) %>%
    ggplot(
        aes(
            x = var,
            y = Estimate
        )
    ) +
    geom_point() +
    geom_errorbar(
        aes(
            ymin = lb,
            ymax = ub
        ),
        color = "red",
        width = .2,
        linetype = "solid",
        linewidth = 0.5
    ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_classic() +
    # rotate x-axis labels 90 degrees
    theme(axis.text.x = element_text(angle = 90)) +
    ylab("WTP") +
    xlab("Main Effects of Carbon Offsetting Attributes")

# consequence message
wtp_tbl%>%
    filter(
        grepl(" consequence", var)
    ) %>%
    ggplot(
        aes(
            x = var,
            y = Estimate
        )
    ) +
    geom_point() +
    geom_errorbar(
        aes(
            ymin = lb,
            ymax = ub
        ),
        color = "red",
        width = .2,
        linetype = "solid",
        linewidth = 0.5
    ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_classic() +
    # rotate x-axis labels 90 degrees
    theme(axis.text.x = element_text(angle = 90)) +
    ylab("WTP") +
    xlab("Interavtive Effects with Consequence Message")

# endorsement
wtp_tbl %>%
    filter(
        grepl(" MetOffice| UN", var)
    ) %>%
    ggplot(
        aes(
            x = var,
            y = Estimate
        )
    ) +
    geom_point() +
    geom_errorbar(
        aes(
            ymin = lb,
            ymax = ub
        ),
        color = "red",
        width = .2,
        linetype = "solid",
        linewidth = 0.5
    ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_classic() +
    # rotate x-axis labels 90 degrees
    theme(axis.text.x = element_text(angle = 90)) +
    ylab("WTP") +
    xlab("Interavtive Effects with Endorsement")

# CO2 consumption
wtp_tbl %>%
    filter(
        grepl(" CO2", var)
    ) %>%
    ggplot(
        aes(
            x = var,
            y = Estimate
        )
    ) +
    geom_point() +
    geom_errorbar(
        aes(
            ymin = lb,
            ymax = ub
        ),
        color = "red",
        width = .2,
        linetype = "solid",
        linewidth = 0.5
    ) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    theme_classic() +
    # rotate x-axis labels 90 degrees
    theme(axis.text.x = element_text(angle = 90)) +
    ylab("WTP") +
    xlab("Interavtive Effects with CO2 Consumption")
```

\begin{itemize}
    \item Higher meeting consumption, i.e., higher co2 value consistently reduce the WTP.
    \item The framing effects is either non-significant or seems to mitigate the preference premium relative to the base line for both endosement and carbon offset approaches.
\end{itemize}

# Latent Class Models, Marginal Utiltiy Without Interaction Terms
This specification allows marginal utility paramters varies by class and 
treating the co2 consumption, framing effect, and PCAs of as demographics.

## Latent Class Treating co2 consumption + framing effect + PCA as demographics, 2 classes
The marginal utility coefficients are:
```{r latent_class_2, echo=FALSE}
summary(lc_1[[1]]$lc_est)
```

The class membership probabilities are:
```{r latent_class_2_class_prob, echo=FALSE}
prob <- apply(lc_1[[1]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_1[[1]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_2_demographics, echo=FALSE}
summary_multinom(lc_1[[1]]$lc_demo)
```

## Latent Class Treating co2 consumption + framing effect + PCA as demographics, 3 classes
The marginal utility coefficients are:
```{r latent_class_3, echo=FALSE}
summary(lc_2[[2]]$lc_est)
```

The class membership probabilities are:
```{r latent_class_3_class_prob, echo=FALSE}
prob <- apply(lc_2[[2]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_2[[2]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_3_demographics, echo=FALSE}
summary_multinom(lc_2[[2]]$lc_demo)
```

\textbf{We can't do an LC esimtation with more than 3 classes due to lack of variation and the hessian matrix would be singular.}

# Latent Class Models, Marginal Utiltiy With Interaction Terms
This specification interacts the marginal utility parameters with the co2 consumption, framing effect, and PCAs.
This specification would be similar to the mixed logit model with the exception that the marginal utility parameters are allowed to vary by class,
but without being random parameters.

## Latent Class interact with co2 consumption + framing effect + PCA, 2 classes
The marginal utility coefficients are:
```{r latent_class_2_int, echo=FALSE}
summary(lc_int_1[[1]]$lc_est)
ic <- c(
    AIC(lc_int_1[[1]]$lc_est),
    BIC(lc_int_1[[1]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_2_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[1]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[1]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_2_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[1]]$lc_demo)
```

### Latent Class interact with co2 consumption + framing effect + PCA, 3 classes
The marginal utility coefficients are:
```{r latent_class_3_int, echo=FALSE}
summary(lc_int_1[[2]]$lc_est)
ic <- c(
    AIC(lc_int_1[[2]]$lc_est),
    BIC(lc_int_1[[2]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_3_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[2]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[2]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_3_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[2]]$lc_demo)
```

### Latent Class interact with co2 consumption + framing effect + PCA, 4 classes
The marginal utility coefficients are:
```{r latent_class_4_int, echo=FALSE}
summary(lc_int_1[[3]]$lc_est)
ic <- c(
    AIC(lc_int_1[[3]]$lc_est),
    BIC(lc_int_1[[3]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_4_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[3]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[3]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_4_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[3]]$lc_demo)
```

\textbf{Based on BIC, the optimal number of classes is 3.
        Also, we could not estiamte a model with more than 4 classes due to lack of variation, i.e., singular hessian matrix.}