---
title: "Consumer WTP for Carbon Offsets"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(data.table)
library(dplyr)
library(stringr)
library(gtools)
library(gmnl)
library(mlogit)
library(ordPens)
library(nnet)
rm(list = ls())
load("/home/yufenlai/Documents/others/carbon_footprint/output/ce_est.RData")

### define a summary function for nnet
summary_multinom <- 
    function(model) {
        s <- summary(model)
        coef_mat <- t(s$coefficients)
        se_mat <- t(s$standard.errors)
        rownames(coef_mat) <- gsub("X", "", rownames(coef_mat))
        rownames(se_mat) <- gsub("X", "", rownames(se_mat))
        
        tbl <-
            lapply(
                1:ncol(coef_mat),
                function(j) {
                    coef <- coef_mat[, j]
                    se <- se_mat[, j]
                    z <- coef / se
                    pval <- 2 * (1 - pnorm(abs(coef / se)))
                    tbl <-
                        data.frame(
                            coef = coef,
                            se = se,
                            z = z,
                            p_val = pval
                        )
                    return(tbl)
                }
            )
        
        names(tbl) <- paste0("Class_", 2:(length(tbl) + 1))
        return(tbl)
    }
```

# Choice Probaility by Information Treatment
```{r choice_prob, echo=FALSE}
data %>%
    mutate(
        opt_out_as_best = as.numeric(alt %in% c(5, 6))
    ) %>%
    group_by(
        framing_effect,
        scenario,
        opt_out_as_best
    ) %>%
    summarise(
        choice_prob = mean(y)
    ) %>%
    pivot_wider(
        names_from = framing_effect,
        values_from = choice_prob
    ) %>%
    filter(opt_out_as_best == 1) %>%
    select(-opt_out_as_best)
```

The above table presents the choice probability of the opt-out option being the best option by information treatment.
The probability differences between different information treatments does not appear to be large.
Evan when the probaiblity of selecting opt out is larger with consequence framing, the difference at about 0.01 percentage points. 

# Summary of PCA
## for Q9
Factor loadings for Q9:
```{r pca_q9, echo=FALSE}
pc_fit_q9_star$pca
```
Importance of components:
```{r pca_q9_importance, echo=FALSE}
summary(pc_fit_q9_star$pca)
```

## for Q10
Factor loadings for Q10:
```{r pca_q10, echo=FALSE}
pc_fit_q10_star$pca
```
Importance of components:
```{r pca_q10_importance, echo=FALSE}
summary(pc_fit_q10_star$pca)
```

# Basic Logit Model
## Basic Logit Model Coefficients
```{r basic_logit, echo=FALSE}
summary(logit_bsc)
```

## Basic Logit Model Willingness to Pay
```{r basic_logit_wtp, echo=FALSE}
logit_bsc$coefficients["price"] <- logit_bsc$coefficients["price"] * -1
wtp.gmnl(logit_bsc, wrt = "price")
```

# Mixed Logit Model
## Mixed Logit Model Coefficients
```{r mixed_logit, echo=FALSE}
summary(mlogit_bsc)
```

## Mixed Logit Model Willingness to Pay
```{r mixed_logit_wtp, echo=FALSE}
mlogit_bsc$coefficients["price"] <- mlogit_bsc$coefficients["price"] * -1
wtp.gmnl(mlogit_bsc, wrt = "price")
```

# mixed logit + co2 consumption + framing effect
## mixed logit + co2 consumption + framing effect Coefficients
```{r mixed_logit_co2, echo=FALSE}
options(width = 300)
summary(mlogit_ctrl_1)
```

## mixed logit + co2 consumption + framing effect Willingness to Pay
```{r mixed_logit_co2_wtp, echo=FALSE}
mlogit_ctrl_1$coefficients["price"] <- mlogit_ctrl_1$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_1, wrt = "price")
```

# mixed logit + co2 consumption + framing effect + PCA
We included the first two coponents for Q9 and Q10.
Q9 was about respondents' attitude towards carbon offsetting,
and Q10 was about respondents' attitude towards climate change.

## mixed logit + co2 consumption + framing effect + PCA Coefficients
```{r mixed_logit_co2_pca, echo=FALSE}
summary(mlogit_ctrl_2)
```

## mixed logit + co2 consumption + framing effect + PCA Willingness to Pay
```{r mixed_logit_co2_pca_wtp, echo=FALSE}
mlogit_ctrl_2$coefficients["price"] <- mlogit_ctrl_2$coefficients["price"] * -1
wtp.gmnl(mlogit_ctrl_2, wrt = "price")
```

\begin{itemize}
    \item Higher meeting consumption, i.e., higher co2 value consistently reduce the WTP.
    \item The framing effects is either non-significant or seems to mitigate the preference premium relative to the base line for both endosement and carbon offset approaches.
\end{itemize}

# Latent Class Models, Marginal Utiltiy Without Interaction Terms
This specification allows marginal utility paramters varies by class and 
treating the co2 consumption, framing effect, and PCAs of as demographics.

## Latent Class Treating co2 consumption + framing effect + PCA as demographics, 2 classes
The marginal utility coefficients are:
```{r latent_class_2, echo=FALSE}
summary(lc_1[[1]]$lc_est)
```

The class membership probabilities are:
```{r latent_class_2_class_prob, echo=FALSE}
prob <- apply(lc_1[[1]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_1[[1]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_2_demographics, echo=FALSE}
summary_multinom(lc_1[[1]]$lc_demo)
```

## Latent Class Treating co2 consumption + framing effect + PCA as demographics, 3 classes
The marginal utility coefficients are:
```{r latent_class_3, echo=FALSE}
summary(lc_2[[2]]$lc_est)
```

The class membership probabilities are:
```{r latent_class_3_class_prob, echo=FALSE}
prob <- apply(lc_2[[2]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_2[[2]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_3_demographics, echo=FALSE}
summary_multinom(lc_2[[2]]$lc_demo)
```

\textbf{We can't do an LC esimtation with more than 3 classes due to lack of variation and the hessian matrix would be singular.}

# Latent Class Models, Marginal Utiltiy With Interaction Terms
This specification interacts the marginal utility parameters with the co2 consumption, framing effect, and PCAs.
This specification would be similar to the mixed logit model with the exception that the marginal utility parameters are allowed to vary by class,
but without being random parameters.

## Latent Class interact with co2 consumption + framing effect + PCA, 2 classes
The marginal utility coefficients are:
```{r latent_class_2_int, echo=FALSE}
summary(lc_int_1[[1]]$lc_est)
ic <- c(
    AIC(lc_int_1[[1]]$lc_est),
    BIC(lc_int_1[[1]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_2_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[1]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[1]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_2_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[1]]$lc_demo)
```

### Latent Class interact with co2 consumption + framing effect + PCA, 3 classes
The marginal utility coefficients are:
```{r latent_class_3_int, echo=FALSE}
summary(lc_int_1[[2]]$lc_est)
ic <- c(
    AIC(lc_int_1[[2]]$lc_est),
    BIC(lc_int_1[[2]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_3_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[2]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[2]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_3_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[2]]$lc_demo)
```

### Latent Class interact with co2 consumption + framing effect + PCA, 4 classes
The marginal utility coefficients are:
```{r latent_class_4_int, echo=FALSE}
summary(lc_int_1[[3]]$lc_est)
ic <- c(
    AIC(lc_int_1[[3]]$lc_est),
    BIC(lc_int_1[[3]]$lc_est)
)
names(ic) <- c("AIC", "BIC")
ic
```

The class membership probabilities are:
```{r latent_class_4_int_class_prob, echo=FALSE}
prob <- apply(lc_int_1[[3]]$lc_est$Wnq, 2, mean)
names(prob) <- paste0("Class_", 1:ncol(lc_int_1[[3]]$lc_est$Wnq))
prob
```

The effects of demographics are:
```{r latent_class_4_int_demographics, echo=FALSE}
summary_multinom(lc_int_1[[3]]$lc_demo)
```

\textbf{Based on BIC, the optimal number of classes is 3.
        Also, we could not estiamte a model with more than 4 classes due to lack of variation, i.e., singular hessian matrix.}